---
title: "Host neighborhood shapes bacterial community assembly and specialization on tree species across a latitudinal gradient"
author: "Genevieve Lajoie"
---


## Libraries

Loading libraries.

```{r Libraries, include=FALSE}
library(reshape2)
library(vegan)
library(ggplot2)
library(dplyr)
library(phytools)
library(Matrix)
library(ape)
library(car)
library(foreach)
library(doParallel)
library(FD)
library(dizzy)
library(nlme)
library(parallel)
library(phyloseq)
library(RColorBrewer)
```

<br />

## Data import and formatting

### Data import

Importing environment with data objects.

General dataset
```{r}
load('~/EcolMon_data.RData')
```

### Study area 

#### Tree species composition

Tree composition : Most abundant at the 25 km scale (across all sites)
```{r}
# Find which 10 species are most abundant across all sites
comm25K.m<-comm25K
comm25K.m$site<-rownames(comm25K.m)
comm25K.m<-melt(comm25K.m, id.vars='site')
ab.sp<-acast(comm25K.m, site~variable, value.var='value')
ab.sum<-sort(colSums(ab.sp), decreasing=T)[1:10]

# Keep only those species
comm25.s<-comm25K.m[which(comm25K.m$variable%in%names(ab.sum)),]

# Plot
ggplot(comm25.s, aes(x="",y=value, fill=variable))+
  geom_bar(stat='identity')+
  scale_fill_brewer(palette = "Paired")+
  coord_polar('y', start=0)+
  theme_bw()+
  facet_wrap(~site)

```

<br />

## Microbial community turnover

#### Dataframes

Define dataframes of microbial and tree community composition
```{r}
## Define mcc matrix
mcc<-decostand(otu_table(ps.epi7K), method='hellinger')

## Host identity per sample
host.id<-data.frame(ps.epi7K@sam_data$Species)
rownames(host.id)<-ps.epi7K@sam_data$Sample_ID
colnames(host.id)<-'Species'

## Host community composition (plot-level)
# Df with one comp. row per sample
Z<-data.frame(sample_data(ps.epi7K))
Z$name<-paste(Z$Site,Z$Type,Z$Plot,sep='-')
comp.host<-comp[match(Z$name,rownames(comp)),]
rownames(comp.host)<-Z$Sample_ID
comp.host<-decostand(comp.host, 'total')
#comp.host<-comp.host[,-which(colSums(comp.host)==0)]

## Host community composition (site-level / 25K)
comp.host.25<-comm25K[match(Z$Site,rownames(comm25K)),]
rownames(comp.host.25)<-Z$Sample_ID

```

Alternative dataframes with ACESAC only
```{r}
# Reformat dataframes: ACESAC only
acesac.samp<-ps.epi7K@sam_data[which(ps.epi7K@sam_data=='ACESAC'),]$Sample_ID
acesac.nm<-which(rownames(mcc)%in%acesac.samp)
comp.host.a<-comp.host[acesac.nm,] # Remove 0-sum columns
comp.host.a<-comp.host.a[,-which(colSums(comp.host.a)==0)]
comp.25.a<-comp.host.25[acesac.nm,]
# comp.25.a<-comp.25.a[,-which(colSums(comp.25.a)==0)] # None
```


Define dataframes of host trait variation (CWM)
```{r}
## Plot-level relative community composition
comp.rel<-decostand(comp, 'total')

# Fill missing trait values with the mean of the species
for (i in 6:9){
  for (j in unique(Z$Species)){
  Z[which(Z$Species==j),i][is.na(Z[which(Z$Species==j),i])==T]<-mean(Z[which(Z$Species==j),i], na.rm=T)
  }
}

# Fill the rest of the values with the mean of everything
for (i in 6:9){
  Z[,i][is.na(Z[,i])==T]<-mean(Z[,i], na.rm=T)
  }

# Check if there are NAs left 
apply(Z[,6:9],2,FUN=function(x) table(is.na(x)))

# Number of traits
trt.str<-data.frame(Z[,which(colnames(Z)%in%c('SLA','Ca','P',"Wood.dens",'name','Species'))])
nt<-ncol(trt.str)-2

# Fill missing values with the mean of the species
for (i in 2:(nt+1)){
  for (j in unique(trt.str$Species)){
  trt.str[which(trt.str$Species==j),i][is.na(trt.str[which(trt.str$Species==j),i])==T]<-mean(trt.str[which(trt.str$Species==j),i], na.rm=T)
  }
}

## Issue with DBH and Height on this -> Should be taking the max, since we had to take samplable small trees sometimes...
## Propagate the highest measure per plot for 

# CWM data frame
cwm.plot<-as.data.frame(matrix(NA, nrow(comp.rel), nt))
rownames(cwm.plot)<-rownames(comp.rel)
colnames(cwm.plot)<-paste(colnames(trt.str)[2:(nt+1)],'cwm', sep='_')

ccm<-NULL # The "mystery" object
# Compute community weighted means
for (j in 2:(nt+1)){
    mat<-acast(name~Species, data=trt.str, value.var=colnames(trt.str)[j]) # Plot x trait matrix
    mat.pa<-decostand(mat, method='pa', na.rm=T) # Presence absence of trait measurements
    new.comp<-comp.rel[,which(colnames(comp.rel)%in%colnames(mat))] # Abundance matrix with species for which there are measurements of this trait
    cwm.plot[,j-1]<-rowSums(mat*new.comp, na.rm=T)/rowSums(mat.pa*new.comp, na.rm=T) # Compute relative abundances only from species where a trait value is reported (for now)
    ccm[[j-1]]<-(new.comp==0)==(is.na(new.comp*mat)) # Output matrix where "F" indicates that there is a trait value missing for a species that is present at a site -> try to find those values / create algorithm to fetch a stand-in value within site or among sites
    rownames(ccm[[j-1]])<-rownames(mat)
}

# Plot names were in the same order throughout the calculations
identical(rownames(mat),rownames(comp.rel))

#######

```

Alternative dataframes with one sample per species per site
```{r}
# Reformat microbial community dataframes: ALL Species
# Keep one occurrence per host species per site (from the plot where it is most abundant)

# Identiying plots where each species is the most abundant within each site
z.comp<-cbind(Z,comp.host)

# Identify sample names of these samples
sam.site<-NULL
for (i in unique(z.comp$Species)){
  for (j in unique(z.comp$Site)){
    sub<-z.comp[which(z.comp$Species==i&z.comp$Site==j),]
    colsp<-which(colnames(sub)==i)
    samid<-sub[sub[,colsp]==max(sub[,colsp]),'Sample_ID']
    sam.site<-c(sam.site,as.character(samid))
  }
}

# Generate new datasets
mcc.sub<-as.data.frame(mcc[which(rownames(mcc)%in%sam.site),])
Z.sub<-Z[which(Z$Sample_ID%in%sam.site),]
comp.host.sub<-comp.host[which(rownames(comp.host)%in%sam.site),]
comp.25.sub<-comp.host.25[which(rownames(comp.host.25)%in%sam.site),]

```

Define host functional diversity at site scale

```{r}
### All traits together
site.list<-list()
site.abund<-list()

# Take the plot where the species is most abundant as the species-site value.
comm25K.b<-comm25K

for (i in unique(Z.sub$Site)){
  # Trait data
  sub<-data.frame(Z.sub[which(Z.sub$Site==i),])
  rownames(sub)<-sub$Species
  sub<-sub[order(rownames(sub)),c(6:9)] # Order the names alphabetically so they will match abund. vector
  site.list[[i]]<-dist(decostand(sub, method='standardize'))
  
  # Abundance data
  abund.vec<-comm25K.b[which(rownames(comm25K.b)==i),]
  abund.vec<-abund.vec/sum(abund.vec)
  abund.vec<-abund.vec[which(names(abund.vec)%in%rownames(sub))]
  abund.vec<-abund.vec[order(names(abund.vec))] # Order the names alphabetically so they will match trait vector
  
  for (j in 1:length(abund.vec)){
    if(abund.vec[j]==0) abund.vec[j]<-0.00000001
  }
  site.abund[[i]]<-abund.vec
 
  print(i)
  print(sum(abund.vec)) # what proportion of the total abundance is captured by the species for which we have trait measurements
}

Fdis.site<-NULL
for (i in 1:length(site.list)){
com.fd<-fdisp(site.list[[i]], as.matrix(site.abund[[i]]))$FDis
Fdis.site<-c(Fdis.site,com.fd)
}


```

Define dataframe of environmental variation
```{r}
## Environmental variation
comp.env<-as.data.frame(Z[,c('MAT','name','Sample_ID')])
rownames(comp.env)<-comp.env$Sample_ID
comp.env<-comp.env[,-which(colnames(comp.env)=='Sample_ID')]
comp.env[,c(1)]<-decostand(comp.env[,c(1)], method='standardize', na.rm=T)
comp.env<-comp.env[order(rownames(comp.env)),]
```

#### Taxonomic description of MCC

```{r}
# By sample
asvphyl<-ps.epi7K@otu_table@.Data
asvphyl<-decostand(asvphyl, 'total')
asvm<-melt(as.matrix(asvphyl))
asvm$Var2<-as.character(asvm$Var2)

taxphyl<-as.data.frame(ps.epi7K@tax_table@.Data)
taxphyl$ASV<-as.character(rownames(taxphyl))

asvt<-merge(asvm, taxphyl, by.x='Var2', by.y="ASV", all.x=T)
asvt<-asvt[-which(asvt$value==0),] # Remove 0 ab lines

# Average across samples
asvt.mean<-aggregate(value~phylum, asvt, sum)
asvt.mean$value<-asvt.mean$value/sum(asvt$value)

ggplot(asvt.mean, aes(x=1,y=value))+
  geom_bar(aes(fill=phylum), colour='black', stat='identity')+
  theme_bw()

# Variability among species
un.s<-ps.epi7K@sam_data$Sample_ID[match(unique(ps.epi7K@sam_data$Species),ps.epi7K@sam_data$Species)]
asvs<-asvt[which(asvt$Var1%in%un.s),]

so<-metadata[which(metadata$Sample_ID%in%un.s),]
so<-so[order(so$Family),]

# Relevel the factors so they are sorted by Tree species family
asvs$Var1<-factor(asvs$Var1, so$Sample_ID[c(29:32,1:6,8:16,23:28,33,7,17:22)]) # 

asvs.a<-aggregate(value~Var1+phylum, asvs, sum)

ggplot(asvs.a, aes(x=Var1, y=value)) +
  geom_bar(aes(fill=phylum), stat='identity', color='black') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size=12),axis.text.y = element_text(size=12)) +
  scale_x_discrete(labels= so$Species[c(29:32,1:6,8:16,23:28,33,7,17:22)]) # 

```

#### MCC ~ Sample structure

All species
```{r}
# Hierarchical plot structure
# PERMANOVA

## All Species - all samples

# TEST OF GENERAL DRIVERS OF COMMUNITY STRUCTURE

# MCC distance matrix
mcc.dist<-vegdist(mcc)
    
# Test for an effect of the structure
plot.str<-data.frame(Z[,which(colnames(Z)%in%c('Species','Plot','Type','Site','MAT'))])
    
# Model
perm.str<-adonis2(mcc.dist ~ Plot:Type:Site + Type:Site + Site, data = plot.str)

# Reporting model
perm.str
    
```

ACESAC only
```{r}
ace.dist<-vegdist(mcc[acesac.nm,])
ace.str<-plot.str[acesac.nm,]
ace.str<-merge(ace.str, metadata[!duplicated(metadata$Site),c('Site','MAT')], by=c('Site'), all.x=T)
    
# Model
perm.ace<-adonis2(ace.dist~ Type:Site + Site , data=ace.str)
# Outputting results
perm.ace
```


### MCC ~ Host Species & Trait Composition

##### All species

Tree taxonomy
```{r}
# Test for an effect of tree taxonomy
tax.str<-data.frame(Z[,which(colnames(Z)%in%c('Family', "Genus", "Species"))])
mod2<-adonis2(mcc.dist ~ Family + Genus:Family + Species:Genus:Family, data = tax.str)
mod2
```

Define model: all species
```{r}
#### Comprehensive model of tree composition across scales

# Variation partitioning (all communities). Using only the most abundant host species for the plot-level and the site-level to reduce collinearity.
mod.comp<-varpart(as.data.frame(mcc), host.id, comp.host[,which(colnames(comp.host)%in%names(ab.sum))], comp.host.25[,which(colnames(comp.host.25)%in%names(ab.sum))], comp.env[,c('MAT')])
plot(mod.comp)

#### Comprehensive model of tree trait composition across scales
# Check correlation between tree traits
trt.str<-data.frame(Z[,which(colnames(Z)%in%c('SLA', 'Ca','P',"Wood.dens"))])
trt.str<-decostand(trt.str, 'standardize')
cor(trt.str)

# Use the full dataset: All species
# Make the full trait dataset
Z.trt<-Z
Z.trt<-merge(Z.trt, decostand(cwm.plot, method='standardize'), by.x=c('name'), by.y=c('row.names'), all.x=T)
Z.trt<-Z.trt[match(rownames(mcc),Z.trt$Sample_ID),]
rownames(Z.trt)<-Z.trt$Sample_ID

### Model per scale

# Host species
tg.sp<-varpart(as.data.frame(mcc), host.id, decostand(Z.trt[,c('SLA', "Wood.dens", "P", "Ca")], method='standardize'), comp.env[,c('MAT')])
plot(tg.sp)

# Host community
tg.plot<-varpart(as.data.frame(mcc),comp.host[,which(colnames(comp.host)%in%names(ab.sum))], comp.env[,c('MAT')])
plot(tg.plot)

# Host site
tg.site<-varpart(as.data.frame(mcc), comp.host.25[,which(colnames(comp.host.25)%in%names(ab.sum))], comp.env[,c('MAT')]) 
plot(tg.site)

```

Plotting results
```{r}
# ALL SPECIES

## HOST SPECIES with individual traits
mcc.dist<-vegdist(mcc, method='bray')
# Pcoa for envfit
mcc.pcoa <- cmdscale(mcc.dist, k=nrow(mcc)-1, eig=T)
# Percent variance per axis
mcc.pcoa$eig/sum(mcc.pcoa$eig)

# Envfit Host species
mi=1
ma=2

# Groups based on host species ID
groups<-host.id
groups<-merge(groups, metadata[,c('Family','Sample_ID')], by.x='row.names', by.y='Sample_ID',all.x=T)
rownames(groups)<-groups$Row.names
groups<-groups[,-1]

# Species with at least 5 observations
absp<-table(groups$Species)[table(groups$Species)>4]

# Envfit of host species ID (only species with at least 5 obs)
efit.host<-envfit(mcc.pcoa, host.id, choices=c(mi:ma))

# Groups based on site identity
gr<-Z[,c('Site','Type','Plot')]
gr$TypePlot<-paste(gr$Type,gr$Plot)

# Envfit of host traits
efit.trt.host<-envfit(mcc.pcoa, decostand(Z.trt[,c('SLA','Ca','P','Wood.dens')], method='standardize'), choices=c(mi:ma), perm=999)

# Envfit of temperature
efit.mat<-envfit(mcc.pcoa, Z.trt[,'MAT', drop=F], choices=c(mi:ma))

# Color vector for families
table(groups$Family) # need 11 colors
colvec<-c(brewer.pal(n = 9, name = "Set1")[-6], brewer.pal(n = 12, name = "Set3")[c(1)], '#FFCC00',brewer.pal(n = 9, name = "Greys")[9])  # remove yellow from Set1
# Reorder colors to attribute red to Sapindaceae and blue to Pinaceae and green to Betulaceae and orange to Fagaceae, Rosaceae in violet
plot(c(1:11), col=colvec, pch=16)
levels(groups$Family)
colvec[c(10,1)]<-colvec[c(1,10)]
colvec[c(7,2)]<-colvec[c(2,7)]
colvec[c(1,3)]<-colvec[c(3,1)]
colvec[c(4,8)]<-colvec[c(8,4)]
colvec.full<-colvec[(as.numeric(groups$Family)-1)] # Remove 1 because n=1 does not exist

# Species-based color vector representing families
groupscol<-groups[order(groups$Species),]
groupscol<-groupscol[!duplicated(groupscol),]
groupscol$Fam<-groupscol$Family
spcol.fam<-as.numeric(groupscol$Family)
spcol.fam<-colvec[(spcol.fam-1)]

# Plotting
# Host species with their traits
par(mar=c(3,3,3,3))
pl<-ordiplot(mcc.pcoa, choices=c(1,2), type='none')
points(pl, 'sites', col=colvec.full, bg=colvec.full, cex=0.8, pch=21) # pch=as.numeric(groups$Species)

# Ellipses
ordiellipse(mcc.pcoa, groups$Species, display = "sites", kind = "se", lty=c(1), conf = 0.95, lwd = 0.25, draw='polygon', col=spcol.fam, alpha = 0.15, choices=c(1,2), show.groups=as.character(names(absp)), label=T) # , label=T

plot(efit.trt.host, p.max=0.05, col='black', cex=1) #col='black',  cex=0.8
plot(efit.mat, lty=3, col='black', cex=1)

# Print leged on the side
legend("center", legend = c(as.character(groupscol$Family[!duplicated(groupscol$Family)])), pch=16, col = c(spcol.fam[!duplicated(spcol.fam)]), cex=1.0)

```

```{r}
### PLOT
## 
mcc.dist<-vegdist(mcc, method='bray')
# Pcoa for envfit
mcc.pcoa <- cmdscale(mcc.dist, k=nrow(mcc)-1, eig=T)

# Envfit Host species
mi=1
ma=2

# Plot composition
mainhost<-names(ab.sum) # OR
mainhost<-names(absp)
efit.plot<-envfit(mcc.pcoa, comp.host[,which(colnames(comp.host)%in%mainhost)], choices=c(mi:ma))

# Color species arrows by family
gc<-groupscol[which(groupscol$Species%in%rownames(efit.plot$vectors$arrows)),]
spcol.fam2<-as.numeric(gc$Family)
spcol.fam2<-colvec[(spcol.fam2-1)]

# MAT
efit.mat<-envfit(mcc.pcoa, Z[,'MAT', drop=F], choices=c(mi:ma))

# Plotting
# Host species with their traits
par(mar=c(3,3,3,3))
pl<-ordiplot(mcc.pcoa, choices=c(1,2), type='none')
points(pl, 'sites', col='grey', bg='grey', cex=0.8, pch=21) # pch=as.numeric(groups$Species)

# Ellipses (do not use species with fewer than 4 observations)

#ordiellipse(mcc.pcoa, groups$Species, display = "sites", kind = "se", lty=c(1), conf = 0.95, lwd = 0.25, draw='polygon', col=spcol.fam, alpha = 0.15, choices=c(1,2), show.groups=as.character(names(absp)), label=T) # , label=T

plot(efit.plot, col=spcol.fam2, cex=0.8) #col='black',  cex=0.8
plot(efit.mat, lty=3, col='black', cex=0.8)

# Output legend separately
legend("topright", legend = c(as.character(gc$Family[!duplicated(gc$Family)])), pch=16, col = c(spcol.fam2[!duplicated(spcol.fam2)]), cex=0.8)

```

```{r}
### Site

## Sites
mcc.dist<-vegdist(mcc, method='bray')
# Pcoa for envfit
mcc.pcoa <- cmdscale(mcc.dist, k=nrow(mcc)-1, eig=T)
# Envfit Host species
mi=1
ma=2

# Plot composition
mainhost1<-names(ab.sum)# Most abundant regionally
mainhost2<-names(absp) # Most abundant in the dataset
mainhost<-intersect(mainhost1,mainhost2) # Intersect between the two
efit.site<-envfit(mcc.pcoa, comp.host.25[,which(colnames(comp.host.25)%in%mainhost)], choices=c(mi:ma), perm=999)

# Color species arrows by family
gc<-groupscol[which(groupscol$Species%in%rownames(efit.site$vectors$arrows)),]
spcol.fam2<-as.numeric(gc$Family)
spcol.fam2<-colvec[(spcol.fam2-1)]

# MAT
efit.mat<-envfit(mcc.pcoa, Z[,'MAT', drop=F], choices=c(mi:ma))

# Ellipses by sites
# Groups based on host species ID
groups.site<-host.id
groups.site<-merge(groups.site, metadata[,c('Site','Sample_ID')], by.x='row.names', by.y='Sample_ID',all.x=T)
rownames(groups.site)<-groups.site$Row.names
groups.site<-groups.site[,-1]


# Plotting
# Host species with their traits
par(mar=c(3,3,3,3))
pl<-ordiplot(mcc.pcoa, choices=c(1,2), type='none')
points(pl, 'sites', col='grey', bg='grey', cex=0.8, pch=21) # pch=as.numeric(groups$Species)

# Ellipses for sites
ordiellipse(mcc.pcoa, groups.site$Site, display = "sites", kind = "se", lty=c(1), conf = 0.95, lwd = 0.25, draw='polygon', alpha = 0.15, choices=c(1,2), label=T) # , label=T


plot(efit.site, p.max=0.05, col=spcol.fam2, cex=0.8) #col='black',  cex=0.8
plot(efit.mat, lty=3, col='black', cex=0.8)

# Add legend afterwards
legend("topright", legend = c(as.character(gc$Family[!duplicated(gc$Family)])), pch=16, col = c(spcol.fam2[!duplicated(spcol.fam2)]), cex=0.8)


```

ACESAC
```{r}
## 
#  Individual-level traits
# Overlay
mcc.dist<-vegdist(mcc[acesac.nm,], method='bray')
# Pcoa for envfit
mcc.pcoa <- cmdscale(mcc.dist, k=nrow(mcc[acesac.nm,])-1, eig=T)
# Percent variance per axis
mcc.pcoa$eig/sum(mcc.pcoa$eig)

# Flip axis 1
mcc.pcoa$points[,1]<-mcc.pcoa$points[,1]*(-1)

# Envfit Host species
mi=1
ma=2

# Envfit of host traits
efit.trt.host<-envfit(mcc.pcoa, decostand(Z.trt[acesac.nm,c(7,8,9,10)], method='standardize'), choices=c(mi:ma), perm=999)

# Envfit of temperature
efit.mat<-envfit(mcc.pcoa, Z.trt[acesac.nm,'MAT', drop=F], choices=c(mi:ma))

# PLOT
# Host species with their traits
par(mar=c(3,3,3,3))
pl<-ordiplot(mcc.pcoa, choices=c(1,2), type='none')
points(pl, 'sites', col='black',  cex=0.8, pch=21) # pch=as.numeric(groups$Species)

plot(efit.trt.host, col='black', cex = 0.8) #, labels = levels(groups$Species)

plot(efit.mat, lty=3, col='black', cex=0.8)

```

```{r}
# Plot-level traits
# Overlay
mcc.dist<-vegdist(mcc[acesac.nm,], method='bray')
# Pcoa for envfit
mcc.pcoa <- cmdscale(mcc.dist, k=nrow(mcc[acesac.nm,])-1, eig=T)

# Flip axis 1
mcc.pcoa$points[,1]<-mcc.pcoa$points[,1]*(-1)

# Envfit Host species
mi=1
ma=2

# Envfit Plot species
mainhost<-names(ab.sum) # or
mainhost<-names(absp)
efit.plot<-envfit(mcc.pcoa, comp.host.a[,which(colnames(comp.host.a)%in%mainhost)], choices=c(mi:ma))

# Color species arrows by family
gc<-groupscol[which(groupscol$Species%in%rownames(efit.plot$vectors$arrows)),]
spcol.fam2<-as.numeric(gc$Family)
spcol.fam2<-colvec[(spcol.fam2-1)]

# Plotting
# Host species with their traits
par(mar=c(3,3,3,3))
pl<-ordiplot(mcc.pcoa, choices=c(1,2), type='none')
points(pl, 'sites', col='black', cex=0.8, pch=21) # pch=as.numeric(groups$Species)

plot(efit.plot, p.max=0.05, col=spcol.fam2, cex=0.8) #col='black',  cex=0.8

legend("topright", legend = c(as.character(gc$Family[!duplicated(gc$Family)])), pch=16, col = c(spcol.fam2[!duplicated(spcol.fam2)]), cex=0.8)


```

```{r}
# SITE
## Sites
mcc.dist<-vegdist(mcc[acesac.nm,], method='bray')
# Pcoa for envfit
mcc.pcoa <- cmdscale(mcc.dist, k=nrow(mcc[acesac.nm,])-1, eig=T)
# Envfit Host species
mi=1
ma=2

# Plot composition
mainhost1<-names(ab.sum)# Most abundant regionally
mainhost2<-names(absp) # Most abundant in the dataset
mainhost<-intersect(mainhost1,mainhost2) # Intersect between the two
efit.site<-envfit(mcc.pcoa, comp.25.a[,which(colnames(comp.25.a)%in%mainhost)], choices=c(mi:ma), perm=999)

# Color species arrows by family
gc<-groupscol[which(groupscol$Species%in%rownames(efit.site$vectors$arrows)),]
spcol.fam2<-as.numeric(gc$Family)
spcol.fam2<-colvec[(spcol.fam2-1)]

# MAT
efit.mat<-envfit(mcc.pcoa, Z[acesac.nm,'MAT', drop=F], choices=c(mi:ma))

# Ellipses by sites
groups.site<-as.data.frame(comp.25.a[,c(1:2)])
groups.site<-merge(groups.site, metadata[,c('Site','Sample_ID')], by.x='row.names', by.y='Sample_ID',all.x=T)
rownames(groups.site)<-groups.site$Row.names
groups.site<-groups.site[,-c(1:3), drop=F]


# Plotting
par(mar=c(3,3,3,3))
pl<-ordiplot(mcc.pcoa, choices=c(1,2), type='none')
points(pl, 'sites', col='black', cex=0.8, pch=21) # pch=as.numeric(groups$Species)

# Ellipses for sites
ordiellipse(mcc.pcoa, groups.site$Site, display = "sites", kind = "se", lty=c(1), conf = 0.95, lwd = 0.25, draw='polygon', alpha = 0.15, choices=c(1,2), label=T) # , label=T


plot(efit.site, p.max=0.05, col=spcol.fam2, cex=0.8) #col='black',  cex=0.8
plot(efit.mat, lty=3, col='black', cex=0.8)

legend("topright", legend = c(as.character(gc$Family[!duplicated(gc$Family)])), pch=16, col = c(spcol.fam2[!duplicated(spcol.fam2)]), cex=0.8)

```



## Specialization

Using the DSI approach: MPD between the interacting partners, compared with randomization of the partner phylogenetic tree (Jorge et al. 2014, Ecol. Let.; Jorge et al. 2017, Oecol.)

#### Site -> species specialization

On 1 species occurrence per site
1- Preparing datasets
```{r}
# DSI as mean phylogenetic distance between hosts

# Use mcc.sub for community composition

# Create a list of dataframes (n=nsite) with host species as rows and ASVs as columns
# Then proceed with calculating the specialization index
# Need count data

# Perform it on the 8 most abundant species at each site
# 8 most abundant host species at every site
# Based on regional abundances 
gb<-Z.trt[which(Z.trt$Sample_ID%in%sam.site),]
gb<-merge(gb,comm25K.m, by.x=c('Site','Species'),by.y=c('site','variable'), all.x=T)
# Some species have 0-values -> change these for the lowest observed across all plots
gb[which(gb$value==0),'value']<-min(comm25K.m[which(comm25K.m$value>0),'value'])

# For each site, calculate the number of host species
table(gb$Site) # -> minimum species is 8

gb.n<-NULL
for (i in 1:length(unique(gb$Site))){
  sub<-gb[which(gb$Site==unique(gb$Site)[i]),]
  sub<-sub[-which(sub$Species=='ACESAC'),]
  sub<-sub[order(sub$value, decreasing=T),]
  gb.n<-rbind(gb.n,sub[c(1:7),])
}

gb.new<-rbind(gb.n,gb[which(gb$Species=='ACESAC'),])

# See if we are still picking the ACESAC at each site (losing 1 - so make sure ACESAC is kept each time)
table(gb.new$Species)
gb.new$Species<-droplevels(gb.new$Species)

# WITH ACESAC = HI
mcc.p<-data.frame(otu_table(ps.epi7K))
mcc.p<-mcc.p[which(rownames(mcc.p)%in%gb.new$Sample_ID),] # gb.new$Sample_ID or sam.site
mcc.p<-mcc.p[,-which(colSums(mcc.p)==0)]
mcc.sub.m<-merge(metadata[,c('Site','Sample_ID','Species')], mcc.p, by.x='Sample_ID', by.y='row.names', all.y=T)

# ASV occurrence data per species and site
mcc.site<-list()
for (i in unique(mcc.sub.m$Site)){
  sub<-mcc.sub.m[which(mcc.sub.m$Site==i),]
  rownames(sub)<-sub$Species
  df.s<-sub[,-which(colnames(sub)%in%c('Species','Site','Sample_ID'))]
  df.s<-df.s[,-which(colSums(df.s)==0)]
  mcc.site[[i]]<-t(df.s)
}

### Pick most abundant ASVs per site

# 500 most abundant
mcc.site.500<-list()
for (i in 1:length(mcc.site)){
  sub<-mcc.site[[i]]
  # Presence of long branches in the phylogeny, remove ASVs from these tips before the selection of the 500 to make sure ASV id numbers are the same between specialization and filtering analyses
  sub<-sub[-which(rownames(sub)%in%longbr.asv),]
  newa<-sort(rowSums(sub), decreasing=T)[1:500]
  mcc.site.500[[i]]<-sub[which(rownames(sub)%in%names(newa)),]
}
names(mcc.site.500)<-names(mcc.site)

```

2- Calculating specialization
```{r}
# Unweighted by host species abundance
# From dataset with 500 ASV most abundant at each site

dsi.site8<-NULL #

for (i in 1:length(mcc.site.500)){

df.site1<-mcc.site.500[[i]] # VAL-HI-P1
# Crop plant phylo
plnt.phy<-drop.tip(comb.tree, comb.tree$tip.label[-which(comb.tree$tip.label%in%colnames(df.site1))])
# All host species weights are equal
comps<-as.data.frame(rep(1,ncol(df.site1)))
rownames(comps)<-colnames(df.site1)

# Run specialization algo
dsi.res<-dsi(df.site1, plnt.phy, comps, Rep = 100, DSICom = T) # Use rep = 1000
# See code for interpretation : https://rdrr.io/github/leorjorge/dizzy/src/R/DSI.R
dsi.res$site<-names(mcc.site.500)[i]
dsi.res$sp<-rownames(dsi.res)

dsi.site8<-rbind(dsi.site8,dsi.res) # or dsi.site.lo

}

hist(dsi.site8$DSI.st)

# Number of ASVs per site (is taking the most abundant 8 actually controlling for site differences in ASV diversity)
lapply(mcc.site.500,nrow)

```

3- Calculating average host species scores
```{r}
# Melting the 500 dataset
# HI
mcc.500.m<-NULL
for (i in 1: length(mcc.site.500)){
  sub<-melt(mcc.site.500[[i]])
  sub$Site<-names(mcc.site.500[i])
  mcc.500.m<-rbind(mcc.500.m,sub)
}

colnames(mcc.500.m)[2]<-c('Species')
mcc.500.m<-merge(mcc.500.m, Z.sub[which(Z.sub$Sample_ID%in%gb.new$Sample_ID),c('Sample_ID','Site','Species')], by=c('Site','Species'), all.x=T)
mcc.500.a<-as.data.frame(acast(mcc.500.m, Sample_ID~Var1, value.var='value', fill=0))

# Community matrix
comm.obs.s<-mcc.500.a

# Merge community composition with the specialization scores
dsi.spe<-merge(mcc.500.m[-which(mcc.500.m$value==0),], dsi.site8, by.x=c('Site','Var1'), by.y=c('site','sp'), all.x=T)

# ASV relative abundances on each plant
comm.sub.m<-melt(as.matrix(comm.obs.s))
comm.sub.m<-merge(comm.sub.m, metadata[,c('Sample_ID','Site','Species')], by.x='Var1',by.y='Sample_ID', all.x=T)
comm.sub.m<-comm.sub.m[-which(comm.sub.m$value==0),]
comm.sub.m<-comm.sub.m[,-1]

# Add ASV abundances to the specialization object
dsi.spe.m<-merge(dsi.spe, comm.sub.m, by.x=c('Site','Species','Var1'), by.y=c('Site','Species','Var2'), all.x=T)
dsi.spe.m$Species<-droplevels(dsi.spe.m$Species)

# Calculate weighted means of ASV specialization scores by host species by site
dsi.spa.w<-NULL
for (i in 1:length(unique(dsi.spe.m$Site))){
  for (j in 1:length(unique(dsi.spe.m$Species))){
    sub<-dsi.spe.m[which(dsi.spe.m$Site==unique(dsi.spe.m$Site)[i]&dsi.spe.m$Species==unique(dsi.spe.m$Species)[j]),]
    if(nrow(sub)>0){
    wgtv<-weighted.mean(sub$DSI.st, sub$value.x)
    dsi.spa.w<-rbind(dsi.spa.w, c(as.character(unique(dsi.spe.m$Site)[i]), as.character(unique(dsi.spe.m$Species)[j]), wgtv))
    }
  }
}

dsi.spa.w<-as.data.frame(dsi.spa.w)
colnames(dsi.spa.w)<-c('Site','Species','DSI.st')
dsi.spa.w$DSI.st<-as.numeric(as.character(dsi.spa.w$DSI.st))

```


#### Explanatory models : Variation among sites

```{r}
### Full object
# Create working object
spec<-dsi.site8
# Add metadata to it
spec<-merge(spec, Z.sub[which(!duplicated(Z.sub$Site)),c('Site','MAT')], by.x=c('site'),by.y=('Site'), all.x=T)
spec$DSI.st[spec$DSI.st<(-1)]<-(-1)

spec$site<-factor(spec$site, levels=c('VAL','VER','SMS','JAC','MEG','GAT','MSH','PRO','FRO'))

# Variation among sites
ggplot(spec, aes(x=site, y=DSI.st))+
  geom_boxplot()+
  theme_bw()+
  theme(axis.text = element_text(size=14), axis.title=element_text(size=16))+
  labs(y='Specialization', x='Site')

# Test
Anova(lm(DSI.st~site, spec), type='III')
TukeyHSD(aov(lm(DSI.st~site, spec)))
```

#### Explanatory models : MAT

```{r}
### CWM object
spec.w<-dsi.spa.w
spec.w<-merge(spec.w, Z.sub[which(Z.sub$Sample_ID%in%gb.new$Sample_ID),], by=c('Site','Species'), all.x=T)

# Model
summary(lm(DSI.st~MAT, spec.w)) # Significant negative relationship


# Only mat
Anova(lm(DSI.st~MAT, spec.w), type='III')

# Distribution of residuals
plot(lm(DSI.st~MAT, spec.w))

# host species and mat
Anova(lm(DSI.st~Species+MAT, spec.w), type='III')

# Distribution of residuals
plot(lm(DSI.st~Species+MAT, spec.w))

```


#### Variation among host species

```{r}

### Does specialization differ among host species?

# Plot this relationship
# Variation observed among species
# Species present in more than 2 plots
nm.sp<-names(table(spec.w$Species)[table(spec.w$Species)>2])

# Family colors

spec.filt3<-spec.w[which(spec.w$Species%in%nm.sp),]
spec.filt3$Species<-droplevels(spec.filt3$Species)
spec.filt3$Species<-factor(spec.filt3$Species, levels=c('ACEPEN','ACERUB','ACESAC','ACESPI','BETALL','BETPAP','FAGGRA','QUERUB','ABIBAL','PICRUB','TSUCAN'))

ggplot(spec.filt3, aes(x=Species, y=DSI.st, fill=Family))+
  geom_boxplot()+
  scale_fill_manual(values=c("#4DAF4A", "#FFCC00", "#377EB8", "#E41A1C"))+
  theme_bw()+
  theme(axis.text = element_text(size=14), axis.title=element_text(size=16), axis.text.x = element_text(angle = 90, hjust = 1), legend.key.size = unit(2,"line"), legend.text=element_text(size=12))+
 # ylim(NA, 1.0)+
  labs(y='Specialization (DSI*)', x='Host species')

# Model
Anova(lm(DSI.st~Family, spec.filt3), type='III')

# Can't do Tukey on Species, not enough samples for all pairwise comparisons. Do it among families.
TukeyHSD(aov(lm(DSI.st~Family, spec.filt3))) # It's alright to use TuKeyHSD for unequal sample sizes because it uses the Tukey-Kramer method which controls for unbalanced designs


```


#### Explanatory models : Host traits

Individual species traits
```{r}
# Plots
ggplot(spec.w, aes(x=SLA, y=DSI.st))+ # Replace SLA by other traits to generate other graphs
  geom_point()+ 
 # geom_smooth(method='lm', se=F)+
  theme_bw()+
  theme(axis.text = element_text(size=16), axis.title=element_text(size=16))+
  labs(y='Specialization', x='Wood dens')

# Models
md.sla<-lm(DSI.st~scale(SLA, center = TRUE, scale = TRUE), spec.w)
summary(md.sla)

md.ca<-lm(DSI.st~scale(Ca, center=T, scale=T), spec.w)
summary(md.ca)

md.p<-lm(DSI.st~scale(P, center=T, scale=T), spec.w)
summary(md.p)

md.wd<-lm(DSI.st~scale(Wood.dens, center=T, scale=T), spec.w)
summary(md.wd)

```

Phylogenetically corrected models

```{r}
# https://lukejharmon.github.io/ilhabela/instruction/2015/07/03/PGLS/
spec.filt<-spec.w

# Need to run with one measure per species
coeftot<-NULL

trt.name<-c('SLA','Ca','P','Wood.dens')

for (j in 1:length(trt.name)){
# Holding coefficient values
coef<-NULL

for (i in 1:1000){
# Generate random pick of one individual per species
newdf <- plyr::ddply(spec.filt, ~Species, function(x){
  ndf <- x[sample(nrow(x),1), ]
})

# Trait A
ssa<-as.data.frame(cbind(scale(newdf[,which(colnames(newdf)==trt.name[j])], center = TRUE, scale = TRUE),newdf$DSI.st))
rownames(ssa)<-newdf$Species
colnames(ssa)<-c('Trt','DSI')

# New tree
plnt2<-drop.tip(comb.tree, comb.tree$tip.label[-which(comb.tree$tip.label%in%rownames(ssa))])
ssa<-ssa[match(plnt2$tip.label,rownames(ssa)),]

# pgls
pglsmodel<-gls(DSI~Trt, data=ssa, correlation=corBrownian(phy = plnt2), method='ML')
anova(pglsmodel)

obj<-c(pglsmodel$coefficients[2], anova(pglsmodel)$'p-value'[2])
obj<-as.numeric(obj)

# Bootstrapping
coef<-rbind(coef, obj)
}
coef<-as.data.frame(coef)
coef$trt<-trt.name[j]

coeftot<-rbind(coeftot, coef)

}

colnames(coeftot)<-c('coef','pval','trt')

for (i in 1:nrow(coeftot)){
  if(coeftot$pval[i]<=0.05) coeftot$pcol[i]<-'green'
  else (coeftot$pcol[i]<-'black')
}

coeftot$trt<-factor(coeftot$trt, c('SLA','Ca','P','Wood.dens'))

ggplot(coeftot, aes(trt, coef))+
  geom_jitter(aes(color=pcol),shape=19, position=position_jitter(0.4))+
  geom_boxplot(alpha=0)+
  theme_bw()+
  theme(axis.text.x = element_text(size=13),axis.text.y = element_text(size=13))


```


#### Explanatory models : Functional diversity 

```{r}
# Merge Fdis metric
spec.filt<-merge(spec.w, Fdis.site, by.x='Site', by.y='row.names', all.x=T)
colnames(spec.filt)[ncol(spec.filt)]<-'fdis.site'

# Plot relationships
ggplot(spec.filt, aes(fdis.site, DSI.st))+
  geom_point()+
  #geom_smooth(method='lm', se=F)+
  theme_bw()+
  theme(axis.text = element_text(size=16), axis.title=element_text(size=16))+
  labs(y='Specialization', x='Host functional diversity')

# Model
summary(lm(DSI.st~fdis.site, spec.filt))

```


#### Explanatory models : Abundance (ACESAC) 

Calculate specialization metric for every ACESAC sample to compare HI and LO abundance
```{r}
# Composition data to work on
mcc.ab<-data.frame(otu_table(ps.epi7K))
# ACESAC samples
Z.trt.a<-Z.trt[which(Z.trt$Species=='ACESAC'),]

# Run in parallel
cl<-makeCluster(7)
registerDoParallel(cl)

dsi.ace8<-foreach(i=unique(gb.n$Site), .combine='rbind', .multicombine=T, .packages=c('dizzy','ape')) %dopar% { #

  # Non ACESAC samples from site i
  samp<-as.character(gb.n[which(gb.n$Site==i), c('Sample_ID')])
  # All ACESAC samples from site i
  samp.a<-as.character(unique(Z.trt.a[which(Z.trt.a$Site==i),c('Sample_ID')]))
  
  # Extract composition data for these samples
  dsi.ace.all<-NULL
  for (j in 1:length(samp.a)){
  # Composition data
  mcc.a<-mcc.ab[which(rownames(mcc.ab)%in%c(samp,samp.a[j])),]
  mcc.a<-mcc.a[,-which(colSums(mcc.a)==0)]
  # Keep only ASV from the 500 at this site
  mcc.a<-mcc.a[,which(colnames(mcc.a)%in%rownames(mcc.site.500[[i]]))]
  mcc.t<-t(mcc.a)
  
  # Add colnames based on host species names
  colnames(mcc.t)<-Z.trt$Species[match(colnames(mcc.t),Z.trt$Sample_ID)]
  
  # Keep only ASV that are observed on ACESAC
  mcc.500.ac<-mcc.t
  # # Keep only those that are observed on ACESAC
  mcc.500.ac<-mcc.t[which(mcc.t[,c('ACESAC')]>0),]
  
  # Calculate specialization of ACESAC ASVs only
  
  # Crop plant phylo
  plnt.phy<-drop.tip(comb.tree, comb.tree$tip.label[-which(comb.tree$tip.label%in%colnames(mcc.500.ac))])
  # All host species weights are equal
  comps<-as.data.frame(rep(1,ncol(mcc.500.ac)))
  rownames(comps)<-colnames(mcc.500.ac)
  
  # Run specialization algo
  dsi.ace<-dsi(mcc.500.ac, plnt.phy, comps, Rep = 100, DSICom = T) # Use Rep = 1000
  # See code for interpretation : https://rdrr.io/github/leorjorge/dizzy/src/R/DSI.R
  dsi.ace$site<-i
  dsi.ace$sp<-rownames(dsi.ace)
  dsi.ace$Sample_ID<-samp.a[j]
  
  dsi.ace.all<-rbind(dsi.ace.all,dsi.ace)
  
  }
  
  dsi.ace.all

}

stopImplicitCluster()

```

Examine relationship with abundance
```{r}
# Merge with plot-level community composition
ab.dsi<-merge(dsi.ace8, comp.host.a[,c('ACESAC'), drop=F], by.x='Sample_ID', by.y='row.names', all.x=T)
colnames(ab.dsi)[ncol(ab.dsi)]<-'ab.local'

# Merge with 25km ACESAC community composition
ab.dsi<-merge(ab.dsi, comp.25.a[,c('ACESAC'), drop=F], by.x='Sample_ID', by.y='row.names', all.x=T)


# Merge with temperature
ab.dsi<-merge(ab.dsi, Z.trt[,c('Sample_ID','MAT')], by='Sample_ID', all.x=T)

# General model

# Number of ASV
length(unique(ab.dsi$sp)) #

# Keep those that occur in all sites to see if there is intra-asv variation in the strength of specialization
# Number of ASV per site # strong gradient variation: maybe control by taking only x asvs per site
sp.site<-aggregate(sp~site, ab.dsi, length)

sp.len<-aggregate(site~sp, ab.dsi, FUN=function(x) length(unique(x)))

ab.dsi.sub<-ab.dsi[which(ab.dsi$sp%in%sp.len$sp[which(sp.len$site==9)]),]
length(unique(ab.dsi.sub$sp)) # 114 present at all sites
# replace the scores less than -1 to -1
ab.dsi.sub$DSI.st[ab.dsi.sub$DSI.st<(-1)]<-(-1)

### Models
# Local abundance
Anova(lm(DSI.st~ab.local*sp, ab.dsi.sub))
plot(lm(DSI.st~ab.local*sp, ab.dsi.sub))


# Regional abundance
Anova(lm(DSI.st~ACESAC*sp, ab.dsi.sub))
plot(lm(DSI.st~ACESAC*sp, ab.dsi.sub))

```

Calculate weighted mean per sample

```{r}
# Community matrix
mcc.ab.m<-mcc.ab[which(rownames(mcc.ab)%in%dsi.ace8$Sample_ID),]
mcc.ab.m<-reshape2::melt(as.matrix(mcc.ab))

# Merge community composition with the specialization scores
dsi.spe.ab<-merge(mcc.ab.m[-which(mcc.ab.m$value==0),], dsi.ace8, by.x=c('Var1','Var2'), by.y=c('Sample_ID','sp'), all.x=T)
dsi.spe.ab<-dsi.spe.ab[complete.cases(dsi.spe.ab),]
dsi.spe.ab$DSI.st[dsi.spe.ab$DSI.st<(-1)]<-(-1)

# Calculate weighted means of ASV specialization scores by host species by site
dsi.spa.ab<-NULL
for (j in 1:length(unique(dsi.spe.ab$Var1))){
    sub<-dsi.spe.ab[which(dsi.spe.ab$Var1==unique(dsi.spe.ab$Var1)[j]),]
    if(nrow(sub)>0){
    wgtv<-weighted.mean(sub$DSI.st, sub$value)
    dsi.spa.ab<-rbind(dsi.spa.ab, c(as.character(unique(sub$site)), as.character(unique(dsi.spe.ab$Var1)[j]), wgtv))
    }
  }

dsi.spa.ab<-as.data.frame(dsi.spa.ab)
colnames(dsi.spa.ab)<-c('Site','Sample_ID','DSI.st')
dsi.spa.ab$DSI.st<-as.numeric(as.character(dsi.spa.ab$DSI.st))


```

Examine relationship with abundance

```{r}

#### Transient object
# spec.ab.old<-spec.ab

# Append the local abundances
spec.ab<-merge(dsi.spa.ab, comp.host.a[,c('ACESAC'), drop=F], by.x='Sample_ID', by.y='row.names', all.x=T)
colnames(spec.ab)[ncol(spec.ab)]<-'ab.local'

# Append the regional abundances
spec.ab<-merge(spec.ab, comp.25.a[,c('ACESAC'), drop=F], by.x='Sample_ID', by.y='row.names', all.x=T)
colnames(spec.ab)[ncol(spec.ab)]<-'ab.reg'


# Plot variation in specialization
# Abundance regional
# Positive relationship
ggplot(spec.ab, aes(x=ab.reg,y=DSI.st))+
  geom_point()+
  geom_smooth(method='lm', se=F)+
  theme_bw()+
  theme(axis.text = element_text(size=16), axis.title=element_text(size=16))+
  labs(y='Specialization', x='Host regional abundance')

# Model
summary(lm(DSI.st~ab.local, spec.ab))

summary(lm(DSI.st~ab.reg, spec.ab))

```
